---
title: "Replication_Assignment"
author: "Anais"
date: "3/14/2022"
output: html_document
---

### Introduction

 Short text description of the study and of the specific data and reanalyses you will be doing, to orient your reader. Outline (briefly!) the goal of the original paper, the data set used, and the analyses conducted, then describe which ones you will replicate.
 
 
 The mayor objectives of this Replication Data Assignment is to replicate the (1) Summary statistics of soil observations (Table II of the paper), (2) the representation of the pH Values of the Morteratsch valley as a function of time since glacial retreat for observations in the fluvial and stable zones (Figure 5. of the paper), and (3) the correlation between the time since glacial retreat and soil properties, for the stable zone of the pro-glacial areas in all three valleys (Table III of the paper).
 In addition, and to train myself working with spatial data, I tried to replicate Figure 3., which shows the soil sample location in the three studied valleys.
 
 
 Therefore, this Rmd document is divided is 4 parts
 
0. Visualization of the 3 Study sites and samples
Slight modification of figure 3 to reproduce the figure with the available data

1. Summary statistic
 Replication of Table II. Summary statistics of soil observations
 
2. Descriptive analysis - Visualization of Figure 5: pH Values as a function of time since glacial retreat for observations in the fluvial and stable zones, with potential debris flows in the stable zone separated from the rest of that zone. 

3.Linear model 
Replication of:
* Table III. Strength (r) and significance (p (r = 0)) of the correlation between the time since glacial retreat and soil properties, for the stable zone of the pro-glacial areas in all three valleys


<!-- 4. Multiregression analysis -->
<!-- Replication of Table IV. Multiple linear regression model performance (r2 value) for soil properties for each valley -->
 
 
### Preliminaries 
-   Install and load in the following libraries:{tidyverse},{rstatix}

```{r}

library(tidyverse)
library(rstatix) #for get_summary_stats()
library(kableExtra) # for re-organizing table (e.g. headers)

```

-   Loading the data

* The soil data set comes from 111 locations visited; 46 in the Morteratsch (M) valley, 36 in the Tschierva (T) valley and 29 in the Forno (F) valley.
Previous analysis were made with an incomplete data set. When I figure it out during spring break, I asked Dr. Temme for another data set. On March 25th he sent me an excel document with many tabs. Then I ordered the data to create a clean CSV to work on R.
I have deleted my previous analysis and present within the document the new R code based on the complete dataset.


```{r}
 f <- "https://raw.githubusercontent.com/anaiszimmer/Data_Analysis_Replication_Assignment/main/Sites.csv"
 
# #f<-"https://raw.githubusercontent.com/anaiszimmer/Data_Analysis_Replication_Assignment/main/Data_Kas.csv" #previous data set
 d <- read_csv(f, col_names = TRUE)
# head(d)
# #View(d)

```
### 0. Study site map - Visualization
```{r}
##Loading additional package for GIS visualization

library(sf) #for ploting shape and coordenates
library(ggplot2)
library(ggspatial) # for annotating maps

## Loading GIS data - glaciers shape - downloaded on GLIMS (https://www.glims.org/maps/glims)
   # I did not wanted to bother Dr.Temme again, so insted of maping the 1850 extent, I used the full glacier data I found online on the GLIMS website.

GLIMSglaciers<-st_read("glims_download_15633/glims_glaciers_polygons.shp")

dgeo <- st_as_sf(d, coords = c("Xdecimal", "Ydecimal"),crs = 4326, agr = "constant")

#Plotting Morteratsch study site
Morteratsch_plots <- ggplot(data = filter(dgeo,Site=="Morteratsch")) +
    geom_sf(data = filter(dgeo,Site=="Morteratsch"), size = 2, shape = 23, fill = "darkred") +
    geom_sf(data = GLIMSglaciers, aes(), color="deepskyblue4", size = 0.2, alpha=0.1) +
    coord_sf(xlim = c(9.860000, 9.980000), ylim = c(46.39000, 46.47000), expand = FALSE)+
  	theme_light() +
	labs(x="Longitude", y="Latitude")+
    annotation_scale(
    location = "tl", width_hint = 0.25,
    pad_x = unit(0.5, "in"),
    pad_y = unit(0.5, "in")) +
  annotation_north_arrow(
    location = "tl", which_north = "true",
    pad_x = unit(0.75, "in"), pad_y = unit(0.8, "in"),
    style = north_arrow_fancy_orienteering)
  #geom_text(data = filter(GLIMSglaciers, glac_name=="MORTERATSCH VADRET DA"), aes(X,Y,label = glac_name), size = 5)
Morteratsch_plots

#Tschierva study site
Tschierva_plots <- ggplot(data = filter(dgeo,Site=="Tschierva")) +
    geom_sf(data = filter(dgeo,Site=="Tschierva"), size = 2, shape = 23, fill = "darkred") +
    geom_sf(data = GLIMSglaciers, aes(), color="deepskyblue4", size = 0.2, alpha=0.1) +
    coord_sf(xlim = c(9.78000, 9.90000), ylim = c(46.40000, 46.45000), expand = FALSE)+
  	theme_light() +
	labs(x="Longitude", y="Latitude")+
    annotation_scale(
    location = "tl", width_hint = 0.25,
    pad_x = unit(0.5, "in"),
    pad_y = unit(0.5, "in")) +
  annotation_north_arrow(
    location = "tl", which_north = "true",
    pad_x = unit(0.75, "in"), pad_y = unit(0.8, "in"),
    style = north_arrow_fancy_orienteering)
Tschierva_plots

#Plotting Forno study site
Forno_plots <- ggplot(data = filter(dgeo,Site=="Forno")) +
    geom_sf(data = filter(dgeo,Site=="Forno"), size = 2, shape = 23, fill = "darkred") +
    geom_sf(data = GLIMSglaciers, aes(), color="deepskyblue4", size = 0.2, alpha=0.1) +
    coord_sf(xlim = c(9.65000, 9.75000), ylim = c(46.32000, 46.38000), expand = FALSE)+
  	theme_light() +
	labs(x="Longitude", y="Latitude")+
    annotation_scale(
    location = "tl", width_hint = 0.25,
    pad_x = unit(0.5, "in"),
    pad_y = unit(0.5, "in")) +
  annotation_north_arrow(
    location = "tl", which_north = "true",
    pad_x = unit(0.75, "in"), pad_y = unit(0.8, "in"),
    style = north_arrow_fancy_orienteering)
Forno_plots
  
par(mfrow=c(3,1))
Morteratsch_plots
Tschierva_plots
Forno_plots

#gridExtra::grid.arrange(Morteratsch_plots,Tschierva_plots,Forno_plots, ncol=3)

```

Previous to this, I had the objective to use a Google map background, as in Figure 3 of the paper. I tried two methods, using the package ggmap, and the package leaflet.
For {ggmap} and get_googlemap(), I learnt that now we need credit on Google cloud to get the necessary API Key. I could have use some of the free credits, but since they ask credit card information and etc. (in case we over passed the number of free credit allowed) I did not want to do it and I look for another method.
Regarding {leaflet} and {sp}, I got some fun interactive maps but it was not what I was looking for. Also I was not able to add the plot points to it.

```{r}

library(sp)
library(leaflet)

df <- data.frame(longitude = runif(10, 9.699000, 9.943500), 
                 latitude = runif(10, 46.33000, 46.45050))

coordinates(df) <- ~longitude+latitude
leaflet(df) %>% addCircleMarkers() %>% addTiles()

```


### 1. Summary statistic
* Table II presents the summary statistics for some numerical attributes of the soils for every pro-glacial valley for the stable and fluvial zones.

```{r}

## computing summary statistic for the 5 variables of interest: pH, SSD, Vegetation cover, Soil Stoniness and Surface Stoniness, for all observations, and for each site depending of the geomorphological zone (Stable or Fluvial)
  
 #reordering Site to read results with the same order than in the paper (easier to check on data values)

d$Site<-factor(d$Site,levels = c("Forno","Tschierva","Morteratsch"))

Summary<-d%>%
  select(Site, geomorph_zone,pH_field, stoniness_soil, stoniness_surf, veg_cover,SSD)%>%
  group_by(Site, geomorph_zone)%>%
  filter(geomorph_zone=="fluvial"| geomorph_zone== "development")%>%
  summarise(across(c("pH_field", "stoniness_soil", "stoniness_surf", "veg_cover","SSD"),list(Mean=mean, SD=sd),na.rm=TRUE)) 
Summary

SummaryAll<-d%>%
  summarise(across(c("pH_field", "stoniness_soil", "stoniness_surf", "veg_cover","SSD"),list(Mean=mean, SD=sd),na.rm=TRUE)) ### !!!!!!!! Need to add a function COUNT that have a na.rm argument
SummaryAll

# Count the number of observation total and per geomorph zone

Count <- d %>%  group_by(Site, geomorph_zone) %>% summarize(n=n())
CountAll<-d %>%filter(geomorph_zone=="fluvial"| geomorph_zone== "development")%>% summarize(n=n())

#joining tables
Summaries<-full_join(SummaryAll,Summary, by=NULL)
Counts<-full_join( CountAll,Count,by=NULL)
table <- left_join(Summaries,Counts,by=(c("Site","geomorph_zone"))) 

table %>%rename(zone='geomorph_zone')%>%
  rename(pH_mean="pH_field_Mean")%>%
  rename(pH_sd="pH_field_SD")%>%
  rename(Number_observations="n")->table

table$Number_observations<-as.factor(table$Number_observations) # needed to transform Number of Obs as a factor if no apllying the transpose t() function will round it at 0.1


table<-table%>%as.data.frame()%>% dplyr::select(Number_observations, pH_mean, pH_sd, SSD_Mean, SSD_SD,veg_cover_Mean,veg_cover_SD, stoniness_soil_Mean, stoniness_soil_SD, stoniness_surf_Mean, stoniness_surf_SD)%>%mutate_if(is.numeric, round, digits=1)%>%t()

#Final table
kableExtra::kbl(table) %>%
  kable_classic()%>%
  kable_paper("striped", full_width = F) %>%
  add_header_above(c("Zone"=1,"All" ,"Stable" = 1, "Fluvial" = 1,"Stable" = 1, "Fluvial" =1 ,"Stable" = 1, "Fluvial" = 1),bold=T) %>%
  add_header_above(c("Pro-glacial area " = 1, "All", "Forno" = 2, "Tschierva" = 2, "Morteratsch" = 2),bold=T)%>%
   pack_rows("Variables",2 ,3)
  

```

* Paper's method: The T-test was used to test for differences in mean soil properties between the different zones in pro-glacial areas.
They tested for pH (pH_field), Depth of soil formation ("SSD" = significant soil depth),Vegetation cover ("veg_cover"), Soil Stoniness ("stoniness_soil"), and Surface stoniness ("stoniness_surf"). The standard deviation was also computed.

```{r}
# Computing the P-Value of the T-test

d<-d%>%filter(geomorph_zone=="fluvial"|geomorph_zone=="development")%>%filter(pH_field!='na')


Table_p_values<-tibble(Site=character(),
                variable=character(),
                p=numeric())

Site<-unique(d$Site)

df<-d%>%select(pH_field, veg_cover, stoniness_soil, stoniness_surf)

Site<-unique(d$Site)
for (i in Site){
   for (j in names(df)){
     #print(j)
 temp<-d %>% filter(Site==i)%>%select(j, geomorph_zone)
 print(nrow(temp))
 p<-t.test(temp[[j]]~temp$geomorph_zone)$p.value
 #print(p)
 r<- tibble(Site=i,variable=j,p=p)
 Table_p_values<-bind_rows(Table_p_values,r)
}}
Table_p_values

Table_p_values<-Table_p_values%>%select(variable,p)%>%
 rename('p-value (stable=fluvial)' ="p")

kableExtra::kbl(Table_p_values) %>%
  kable_classic()%>%
  kable_paper("striped", full_width = F) %>%
  pack_rows("Morteratsch",1 ,4)%>%
  pack_rows("Tschierva",5 ,8)%>%
  pack_rows("Forno",9 ,12)

```

### 2. Descriptive analysis - Visualization of Figure 5: pH Values as a function of time since glacial retreat for observations in the fluvial and stable zones, with potential debris flows in the stable zone separated from the rest of that zone. 

```{r, include=TRUE}

# Creating a new data set to remove "slope" and "none" values of the [geomorh_zone variable]
d_stability_Forno<-d%>%filter(Site=="Forno")%>% ###COMMENTS : It took me a very long while to understand the graph was only including the Forno data set!
  filter(geomorph_zone!="none")%>%
  filter(geomorph_zone!="slope")

# Creating a new variable call stability
d_stability_Forno<-d_stability_Forno%>%
  mutate(stability=case_when(
    geomorph_zone=="development" & potential_dist=="debris flow"~"stable-debris",
    geomorph_zone=="development"~"stable",
    geomorph_zone=="fluvial"~"fluvial"))

#View(d_stability_Forno)

# Reordering the stability variable
d_stability_Forno$stability<-factor(d_stability_Forno$stability,levels=c("stable-debris", "stable","fluvial"))

# Plotting pH in function of time since glacier retreat , and grouping by stability

pH_time<-ggplot(d_stability_Forno, aes(y=pH_field, x=realTSR, color=stability)) +
  geom_point(aes(color=stability), size=3) +
  scale_color_manual(values=c('hotpink3','darkgreen','deepskyblue3'))+
  theme_bw()+ 
  coord_cartesian(xlim=c(0,120),ylim=c(4,8)) +
  scale_x_continuous(breaks = c(0, 20, 40, 60, 80, 100,120))+
  labs(x="time since glacial retreat (y)", y="pH") +
  theme_bw()
pH_time

```
* On this first version of the graph we observe three differences with the original graph of the paper. One fluvial point is missing from the data base and one points (top right) appear as stable and should be a fluvial point, as weel its neighboor that is also stable and should be stable-debris.


Below we look at the same graph but using different shapes to draw the pH values. Here, we notice that F27 and F29 where superposed. 
The blue fluvial F27 was hidden by the green stable 29.


```{r}
pH_time<-ggplot(d_stability_Forno, aes(y=pH_field, x=realTSR, color=stability)) +
  geom_point(aes(shape=stability,color=stability), size=3) +
  scale_color_manual(values=c('hotpink3','darkgreen','deepskyblue3'))+
  scale_shape_manual(values=c(22,21,24))+
  theme_bw()+ 
  coord_cartesian(xlim=c(0,120),ylim=c(4,8)) +
  scale_x_continuous(breaks = c(0, 20, 40, 60, 80, 100,120))+
  labs(x="time since glacial retreat (y)", y="pH") +
  ggrepel::geom_text_repel(data=d_stability_Forno,mapping= aes(label=d_stability_Forno$point),size=3,box.padding=unit(0.5,"lines"))+
  theme_bw()
pH_time

```

Similar conclusion are observed using geom_jitter(), and playing with the transparency alpha:

```{r}

pH_time<-ggplot(d_stability_Forno, aes(y=pH_field, x=realTSR, color=stability)) +
  geom_jitter(aes(color=stability), size=3, alpha=0.5) +
  scale_color_manual(values=c('hotpink3','darkgreen','deepskyblue3'))+
  theme_bw()+ 
  coord_cartesian(xlim=c(0,120),ylim=c(4,8)) +
  scale_x_continuous(breaks = c(0, 20, 40, 60, 80, 100,120))+
  labs(x="time since glacial retreat (y)", y="pH") +
  ggrepel::geom_text_repel(data=d_stability_Forno,mapping= aes(label=d_stability_Forno$point),size=3,box.padding=unit(0.5,"lines"))+
  theme_bw()
pH_time

```

However, there is still a difference according F28 that appear as stable-debris (red)  on the original figure in Figure 5, and as stable (green) in our graph.
And one other fluvial point between F24 and F22 is missing.

Looking again a the dataset shared by the authors, there is another dataset tab, called All_export. The data of this data set is identic to the Site dataset, but except for the variable Potential disturbance where the point F28 is considered as debris flow instead of NA.

Below the same code is run with this alternative data set called "All Export".

This disagreement might be what also cause the different results observed in the Summary table.But we have the same count of observation, so this is still questionable.

Now only an error with F23 which appear with "slope" in the both dataset, but as "fluvial" in the graph of the paper.

```{r}

All_export <- read_csv("All_export.csv")

d_stability_Forno2<-All_export%>%
  filter(Site=="Forno")%>%
  mutate(stability=case_when(
    geomorph_zone=="development" & potential_dist=="debris flow"~"stable-debris",
    geomorph_zone=="development"~"stable",
    geomorph_zone=="fluvial"~"fluvial"))%>%
  filter(pH_field!="na")%>%
  filter(geomorph_zone!="na")

d_stability_Forno2$stability<-factor(d_stability_Forno2$stability,levels=c("stable-debris", "stable","fluvial"))

pH_time2<-ggplot(d_stability_Forno2, aes(y=pH_field, x=realTSR, color=stability)) +
  geom_jitter(aes(color=stability), size=3, alpha=0.5) +
  scale_color_manual(values=c('hotpink3','darkgreen','deepskyblue3'))+
  theme_bw()+ 
  coord_cartesian(xlim=c(0,120),ylim=c(4,8)) +
  #scale_x_continuous(breaks = c(0, 20, 40, 60, 80, 100,120))+
  labs(title="All export data set", x="time since glacial retreat (y)", y="pH") +
  ggrepel::geom_text_repel(data=d_stability_Forno2,mapping= aes(label=d_stability_Forno2$point),size=3,box.padding=unit(0.5,"lines"))+
  theme_bw()
pH_time2

```

### 3. Linear model


Linear models for the various numerical soil properties were estimated with time and topographical properties as explanatory variables in order to assess the joint effect of time and topography.
Table IV summarizes model performance for each of the soil properties in each of the valleys.

* Table III. Strength (r) and significance (p (r = 0)) of the correlation between the time since glacial retreat and soil properties, for the stable zone of the
pro-glacial areas in all three valleys

```{r}



f <- "https://raw.githubusercontent.com/anaiszimmer/Data_Analysis_Replication_Assignment/main/Sites.csv"

#f<-"https://raw.githubusercontent.com/anaiszimmer/Data_Analysis_Replication_Assignment/main/Data_Kas.csv" #previous data set
d <- read_csv(f, col_names = TRUE)
head(d)
#View(d)

 # Creating a new data set to remove "slope" and "none" values of the [geomorh_zone variable]
d_stability<-d%>% 
  filter(geomorph_zone!="none")%>%
  filter(geomorph_zone!="slope")

 
 # Creating a new variable call stability
d_stability<-d_stability%>%
  mutate(stability=case_when(
    geomorph_zone=="development" & potential_dist=="debris flow"~"stable-debris",
    geomorph_zone=="development"~"stable",
    geomorph_zone=="fluvial"~"fluvial"))

d_stability$potential_dist[is.na(d_stability$potential_dist)]<-"none"

 
### Pearson Correlation 

#Subset stable zone only
dcor<-d_stability%>%filter(stability=="stable")%>%filter(potential_dist=="none")%>%select(Site,pH_field, SSD,veg_cover, stoniness_soil, stoniness_surf, realTSR)%>%filter(!is.na(pH_field))

dcor$Site<-factor(dcor$Site,levels = c("Forno","Tschierva","Morteratsch"))
# dcor<-d%>%filter(geomorph_zone=="development")%>%filter(potential_dist!="debris flow")%>% select(Site,pH_field, SSD,veg_cover, stoniness_soil, stoniness_surf, realTSR)%>%filter(!is.na(pH_field))
# view(dcor)

Site<-unique(dcor$Site)

dvar<-dcor%>%select(pH_field,SSD, veg_cover, stoniness_soil, stoniness_surf)



#Empty table for results
Table_pearson<-tibble(Site=character(),
                       variable=character(),
                       r=numeric(),
                       p=numeric())

for (i in Site){
  for (j in names(dvar)){
    #print(j)
    temp<-dcor %>% filter(Site==i)
    print(nrow(temp))
    p<-cor.test(temp[[j]],temp$realTSR,method="pearson")$p.value
    rho<-cor.test(temp[[j]],temp$realTSR,method="pearson")$estimate
    #print(p)
    results<- tibble(Site=i,variable=j,r=rho,p=p)
    Table_pearson<-bind_rows(Table_pearson,results)
  }}
Table_pearson

Table_pearson<-Table_pearson%>%select(variable,r,p) #remove the column site for nicer table in kableExtra
(count <- dcor %>%  group_by(Site) %>% summarize(n=n()))    # Count the number of observation total and per geomorph zone

  kableExtra::kbl(Table_pearson) %>%
  kable_classic()%>%
  kable_paper("striped", full_width = F) %>%
  pack_rows("Morteratsch  n=31",1 ,5)%>%
  pack_rows("Tschierva  n=29",6 ,10)%>%
  pack_rows("Forno  n=22",11 ,15)

 
```

The results obtained are very different than the one of Table III. The sign of r is most of the time the same than show in Table III (except for pH at Forno).
At the Forno, the authors used 20 points, I found 22 points in the table that have stable carracteristic and no NA values. It seems two points have be deleted from the data set. I cannot find which ones.


The same analysis were ran using the All export data (in can different results might be observed, more similar to the paper) and not difference were noted.

